\text{Additive drift with tail bounds}

### 2021-10-13 article.tex
add some fixes

\begin{theorem}[Drift Theorem for Lower Bounds] \label{thm:neat}
    Let $\{X_t\}_{t \ge 0}$ be a Markow process over a finite set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers. Pick two real numbers $a$ and $b$ such that $a < b$ and let the random variable $T$ denote
    the earliest point in time $t \ge 0$ where $\mathbf{g}(X_t) \leq a$ holds.

    If there are constants $\lambda > 0$, $D \ge 1$, and a $p > 1$ taking only positive values, for which the following four conditions hold
    \renewcommand\labelenumi{(\theenumi)}
    \begin{enumerate}
        \item $\gfun(X_0) \ge b$,
        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - \gfun(X_t)))} |\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} =: \rho$,
        \item $\forall t \ge 0 E\left[e^{-\lambda(\gfun(X_{t + 1} - b))} |\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
    \end{enumerate}
    then for all time bounds $B \ge 0$ the probability that $T$ exceeds $B$ is at most

@@

\begin{lemma}[Lemma 2.8 in \cite{Hajek 82}] \label{lm:haj}
    If conditions 2 and 3 are met then for all $t \geq 0$ we have

@@

\end{proof}
\subsection{Theorem for sub-gaussian processes}
\cm
\subsection{Comparing requirements}
\label{sec:compare}
% To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
% $$
%     \exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] $$$$
%     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$

@@

To compare rigors of requirements we need to explore behavior $E[e^{\gamma X}]$ depending to $\gamma$. This behavior depends on the distribution $X$, hence we consider the two cases for continuos and for discrete distributions.

@@

By the definition of expectation of a function we have
\[
    E[e^{\gamma X}] = \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \,dx,
\]

@@

We compute the first derivative of this expectation over $\gamma$

@@

We denote $c = \int_{0}^{+\infty} x^2 f(x) \, dx \geq 0$.If we integrate the inequality above, we obtain a lower bound on the expectation.

\[
    E[e^{\gamma X}] = E[1] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}.
\]

@@

In the same way we consider the expectation of function by definition

@@

Denote set $P = \{i \in \mathbb{N}: x_i \geq 0\}$. Hence

@@

When we take the first derivative we can split to two group, first of which are terms with indices from $P$ set and others
% В таком случае, аналогично возьмем производную и рассмотрим пот отдельности слагаемые из множества P и остальные:

@@

Since the bounds for both cases are the same, we do not distinguish cases for continuos and discrete v.v.

@@

In this case, the expectation of the exponent cannot be lower than 1, and also in the some neighborhood of zero it grows faster than any $e^{\frac{\gamma^2 c}{2}}$, since

@@

In this section we consider some specific family of continuos distributions with symmetric probability density functions, because their expectation equals 0.

@@

Consider function

$$f(x) = c \expx{a},\ a > 0.$$

@@

\begin{align*}
    \infint{f(x)} & = \infint{c\expx{a}} = 1,
\end{align*}

\begin{align*}
    \infint{\expx{a}} = 2\infintpoz{\expxpoz{a}}.
\end{align*}

@@

Then we can consider $e^{\gamma X}$ as a infinite series

@@

Where by $a_n$ we denote $\frac{E[X^{2n}]n!}{2n!}$ for all $n \in \mathbb{N}$.

Third equality is satisfied because all odd momentums are 0 due to the symmetry of the function.

@@

We now consider two cases depending on the value of $a$
\begin{caseof}
    \case{$a \geq 2$}{

        Then $A$ exist, which implies that there exist $C > 0$ such that
        $\forall n \geq 0 a_n \leq C^n$.
        Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}.$
    }
    \case{$a < 2$}{

@@

We can compute same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}_{n \in \mathbb{N}}$ and threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.

In this case we need to explore behavior of series when $a \in (1, 2)$. By another definition of sub-gaussian process

>>>>

\begin{theorem}[Drift Theorem for Lower Bounds] \label{thm:neat}
    Let $\{X_t\}_{t \ge 0}$ be a Markow process over a finite set of states $\mathbb{S}$, and $\mathbf{g} : \mathbb{S} \rightarrow \mathbb{R}$ a function that assigns to every state a non-negative real numbers. Let $a$ and $b$ be some real numbers such that $a < b$ and let the random variable $T$ denote
    the earliest point in time $t$ when $\mathbf{g}(X_t) \leq a$ holds.

    If there are constants $\lambda > 0$, $D \ge 1$, and $p > 1$ such that the following three conditions hold
    \renewcommand\labelenumi{(\theenumi)}
    \begin{enumerate}
        \item $\gfun(X_0) \ge b$,
        \item for all \(t \geq 0\) $ E\left[e^{-\lambda(\gfun(X_{t + 1}) - \gfun(X_t))} \Big|\ X_t,\ \gfun(X_t) < b\right] \leq 1 - \frac{1}{p} =: \rho$ and
        \item for all \(t \geq 0\) $ E\left[e^{-\lambda(\gfun(X_{t + 1}) - b)} \Big|\ X_t,\ \gfun(X_t) \ge b\right] \leq D$,
    \end{enumerate}
    then for all time bounds $B \ge 0$ the probability that $T$ exceeds $B$ is at most

@@

\begin{lemma}[Lemma 2.8 in \cite{Hajek82}] \label{lm:haj}
    If conditions 2 and 3 are met then for all $t \geq 0$ we have

@@

\end{proof}
\subsection{Theorem for sub-gaussian processes}

\begin{theorem}
    Suppose $X_t$ has a drift of at most $\varepsilon > 0$ and $(X_t - \varepsilon t)$ is $(c, \delta)$ - sub-Gaussian. Let \(X_0\) be \(a \in \mathbb{R}\) and let \(T\) be the first moment in time when \(X_t\) exceeds some \(b > a\). Then for all \(s \leq \frac{b - a}{2\varepsilon}\) we have

    \[\Pr[T \leq s] \leq exp\left(-\frac{b - a}{4} min\left(\delta, \frac{b - a}{2cs}\right)\right)\]
\end{theorem}

\subsection{Comparing requirements}
\label{sec:compare}
% To use some versions of drift theorems, we need some requirements on the sequence of random variables to be satisfied, such as the following:
% $$
%     \exists \delta \geq 0,\ c \in \mathbb{R} \ \forall  t \geq 0 \ \forall \gamma \in [0, \delta] $$$$
%     E\left[e^{\gamma (X_{t + 1} - X_{t})}\right] \leq e^{\frac{c \gamma^2}{2}},$$

@@

To compare rigors of requirements we need to explore behavior of $E[e^{\gamma X}]$ depending to $\gamma$. This behavior depends on the distribution $X$, hence we consider the two cases for continuos and for discrete distributions.

@@

By the definition of the expectation of a function we have
\[
    E[e^{\gamma X}] = \int_{-\infty}^{+\infty} f(x) e^{\gamma x} \,dx,
\]

@@

Since we assume thet the integral is converging, compute the first derivative of this expectation over $\gamma$

@@

We denote $c = \int_{0}^{+\infty} x^2 f(x) \, dx \geq 0$. If we integrate the inequality above, we obtain a lower bound on the expectation.

\[
    E[e^{\gamma X}] = E[e^{0}] + \int_0^\gamma \frac{\partial E[e^{\gamma X}]}{\partial \gamma} \, d\gamma \geq 1 + \gamma E[X] + \gamma^2 \frac{c}{2}.
\]

@@

We now show a similar result for a discrete r.v. X. Consider the expectation of function by definition

@@

Let $P$ be the set of indices of non-negative values of X that is \(P = \{i \in \mathbb{N} : x_i \geq 0\}\). Hence

@@

Therefore, we compute the derivative of \(E\left[e^{\gamma X}\right]\)

@@

We obtained a similar lower bound, apart from the constant \(c\). Hence,
further in this section we do not specify if consider a process has a continuous or discrete distribution. Note also that in both cases \(c\) is bounded by the second momentum of \(X\).

@@

When we have \(E\left[X\right] > 0\), the expectation of the exponent cannot be lower than 1, and also in the some neighborhood of zero it grows faster than any $e^{\frac{\gamma^2 c}{2}}$, since

@@

We consider some specific family of continuos distributions with symmetric probability density functions with zero expectation.

@@

Consider random variable \(X_a\) with probability density

$$f(x) = c \expx{a},$$

where \(a > 0\) is a parameter of the distribution.

@@

\begin{align*}
    \infint{f(x)} & = \infint{c\expx{a}} = 1, \\
    \infint{\expx{a}} & = 2\infintpoz{\expxpoz{a}}.
\end{align*}

@@

Note that all odd momentums are zero.

Now we consider $e^{\gamma X}$ as a infinite series.

@@

where by $a_n$ we denote $\frac{E[X^{2n}]n!}{2n!}$ for all $n \in \mathbb{N}$.

@@

We now consider two cases depending on the value of the parameter $a$ of r.v. \(X\).

\begin{caseof}
    \case{$a \geq 2$}{

        Then $A$ is finite, which implies that there exist $C > 0$ such that
        for all $ n \geq 0 $ we have $a_n \leq C^n$.
        Hence $E[e^{\gamma X}] \leq e^{\gamma^2 C}$, because

        \[
            E[e^{\gamma X}] = \sum_{n = 0}^{+\infty}\frac{(\gamma^2)^{n}}{n!} a_n \leq \sum_{n = 0}^{+\infty}\frac{(\gamma^2 C)^{n}}{n!} = e^{\gamma^2 C}
        \]
    }
    \case{$a < 2$}{

@@

We can compute the same limit for $\{\sqrt[n]{\frac{a_n}{n!}}\}_{n \in \mathbb{N}}$ and the threshold regarding to $a$ is 1, where only for $a \geq 1$ series converge.

In this case we need to study the behavior of series when $a \in (1, 2)$. By another definition of sub-gaussian process
